Case-Base Neural Network: survival analysis with time-varying, higher-order interactions

### 2023-06-30 manuscript/manuscript.tex
updated manuscript machine learning with applicat

\documentclass[APA,LATO1COL]{CBNNManuscript}
\usepackage{lmodern}

\articletype{Article Type}%

\received{XXXXXX}
\revised{XXXXXX}
\accepted{XXXXXX}

\raggedbottom

@@

@@

\author[1]{Jesse Islam}

\author[2]{Maxime Turgeon}

\author[1,3]{Robert Sladek}

\author[4]{Sahir Bhatnagar}

\address[1]{\orgdiv{Department of Quantitative Life Sciences,} \orgname{McGill University}, \orgaddress{\state{Quebec}, \country{Canada}}}

\address[2]{\orgdiv{Department of Statistics}, \orgname{University of Manitoba}, \orgaddress{\state{Manitoba}, \country{Canada}}}

\address[3]{\orgdiv{Department of Human Genetics}, \orgname{McGill University}, \orgaddress{\state{Quebec}, \country{Canada}}}

\address[4]{\orgdiv{Department of Biostatistics}, \orgname{McGill University}, \orgaddress{\state{Quebec}, \country{Canada}}}

\corres{Jesse Islam \email{jesse.islam@mail.mcgill.ca}}

\presentaddress{McGill Genome Center: 740 Dr Penfield Ave, Montreal, Quebec H3A }

\abstract[Summary]{Neural network-based survival methods can model data-driven covariate interactions. While these methods can provide better predictive performance than regression-based approaches, not all can model time-varying interactions and complex baseline hazards. To address this, we propose Case-Base Neural Networks (CBNNs) as a new approach that combines the case-base sampling framework with flexible neural network architectures. Using a novel sampling scheme and data augmentation to naturally account for censoring, we construct a feed-forward neural network that may take time as an input. CBNNs predict the probability of an event occurring at a given moment to estimate the hazard function. We compare the performance of CBNNs to regression and neural network-based survival methods in a simulation and three case studies using two time-dependent metrics. First, we examine performance on a simulation involving a complex baseline hazard and time-varying interactions to assess all methods, with CBNN outperforming competitors. Then, we apply all methods to three real data applications, with CBNNs outperforming the competing models in two studies and showing similar performance in the third. Our results highlight the benefit of combining case-base sampling with deep learning to provide a simple and flexible modeling framework for data-driven, time-varying interaction modeling of single event survival outcomes. An R package is available at https://github.com/Jesse-Islam/cbnn.}

\keywords{survival analysis, machine learning, case-base, neural network}

\maketitle

@@

Neural networks provide a data-driven approach to approximating interaction terms that benefit prediction. For example,
DeepSurv is a neural network-based proportional hazards model that implements the Cox partial log-likelihood
as a custom loss function \citep{katzman2018DeepSurv}. This results in a stepwise absolute risk curve that cannot
accommodate time-varying interactions. DeepSurv has shown better performance than Cox proportional hazard models in the Rotterdam and German Breast Cancer Study Group and The Molecular Taxonomy of Breast Cancer International Consortium datasets \citep{katzman2018DeepSurv}. It has also been used to develop personalized prediction models for cancers in a number of studies \citep{ds1} \citep{ds2} \citep{ds3}. Previous work suggests we may change the loss function to address
non-proportional hazards \citep{faraggi1995neural}. In contrast, DeepHit assumes an inverse Gaussian distribution
as the baseline hazard \citep{lee2018DeepHit} and specifies each follow-up time of interest in the model. DeepHit
directly estimates survival curves, rather than deriving a hazard function \citep{lee2018DeepHit}. On real datasets,
DeepHit outperformed DeepSurv \citep{lee2018DeepHit}. We note that DeepSurv and DeepHit both use custom loss
functions \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}. In addition, DeepHit introduces a hyperparameter
weighing its two loss functions (negative log-likelihood and ranking losses) \citep{lee2018DeepHit}. Compared
to these neural network methods, regression-based approaches require prior specification of all interaction terms,
which makes it challenging to model covariate effects that change with time. The current neural network models
provide flexibility at the cost of clarity, while regression models provide clarity at the cost of flexibility.

@@

{\centering \includegraphics[width=1\linewidth]{../figures/Figure1.pdf}}

@@

To choose our method-specific hyperparameters, we use the Integrated Brier Score (IBS) \citep{graf1999}, which is based on the
Brier Score (BS) and provides a summarized assessment of performance for each model. We assess the performance of models on
a held out dataset using two metrics: 1) the Index of Prediction Accuracy (IPA) \citep{kattan2018index}; and 2) the Inverse probability
censoring weights-adjusted time dependent area under the receiver operating characteristic curve ($AUC_{IPCW}$) \citep{auc}. The
IPA score is a metric for both discrimination and calibration, while the $AUC_{IPCW}$ provides a metric for discrimination only.

@@

The IPA is a function of time, based on the BS. The IPA score is given by \begin{align}
\textrm{IPA}(t) &= 1-\frac{BS_{model}(t)}{BS_{null}(t)}, \nonumber
\end{align} where \(BS_{model}(t)\) represents the BS over time \(t\) for the model of interest and \(BS_{null}(t)\) represents
the BS if we use an unadjusted Kaplan-Meier (KM) curve as the prediction for all observations \citep{kattan2018index}. The
IPA score has an upper bound of one, where positive values show an increase in performance over the null model and
negative values show that the null model performs better. As an extension of BS, the IPA score assesses both model calibration
and discrimination and show how performance changes over follow-up time \citep{graf1999} \citep{kattan2018index}.

\hypertarget{auc}{%
\subsubsection{Inverse probability censoring weights-adjusted time dependent area under the receiver operating characteristic curve ($AUC_{IPCW}$) }\label{auc}}

The IPCW-adjusted AUC ($AUC_{IPCW}$) is a time-dependent metric that considers censoring \citep{auc}. For a given follow-up time of interest,
 \begin{align*}
AUC_{IPCW}(t)=\frac{\sum^{n}_{i=1}\sum^{n}_{j=1} I \left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right) \cdot I(T_{i}\leq t,\delta_{i}=1) \cdot \left( 1-I(T_{j}\leq t,\delta_{j}=1) \right) \cdot W_{i}(t) \cdot W_{j}(t)} { \sum^{n}_{i=1}\sum^{n}_{j=1} I(T_{i}\leq t,\delta_{i}=1) \cdot \left( 1-I(T_{j}\leq t,\delta_{j}=1) \right) \cdot W_{i}(t) \cdot W_{j}(t) },
\end{align*}where \begin{align*}
W_{i}(t)=\frac{I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} +\frac{I(T_{i}>t)}{\widehat{G}(t)}.
\end{align*} The $AUC_{IPCW}$ measures discrimination \citep{auc}. By examining both IPA and $AUC_{IPCW}$,
we can better understand whether a model performs better in terms of calibration or discrimination.

@@

Neural networks require external parameters that constrain the learning process (hyperparameters). We apply the following hyperparameter
optimization procedure to each method for each study. First, we fix a test set with 15\% of the data which we keep aside during hyperparameter selection.
To determine the best hyperparameters for each neural network method, we use a three-fold cross-validated grid search \citep{gulli2017} on the remaining
data (85\% training, 15\% validation) for the following range of hyperparameters:
\[
Grid\ search: \begin{cases}
Learning\ rate \sim \{0.001, 0.01\} \\ 
Dropout \sim \{0.01,0.05, 0.1\} \\
First\ layer\ nodes \sim \{50, 75, 100\} \\
Second\ layer\ nodes \sim \{10,25,50\} \\
Number\ of\ batches \sim \{100, 500\} \\
Activation\ function \sim \{ReLU, Linear\}\\
\alpha \sim \{0, 0.5, 1\}. \ (DeepHit \ only)
\end{cases}
\]

\begin{table}
\caption{Hyperparameters selected after three-fold cross-validated grid search along with the average IBS for each neural network model in the
complex simulation (A); multiple myeloma (MM) case study (B); free light chain (FLC) case study (C); and prostate cancer (Prostate) case study (D).}
\label{tab:wins}
\begin{center}\includegraphics[width=1\linewidth]{../figures/Table1.pdf} \end{center}
\end{table}

@@

@@

{\centering \includegraphics[width=1\linewidth]{../figures/Figure2.pdf}

@@

We simulate data to evaluate the performance of CBNN in comparison with existing regression (Cox, CBLR) and neural network (DeepHit, DeepSurv)
methods. We specify a linear combination of each covariate as the linear predictor in the regression-based methods (Cox, CBLR), which contrasts
with neural network approaches that allow for approximations of non-linear interactions. The simulated data are based on a complex baseline hazard with time-varying
interactions, each with 10\% random censoring. We simulate three covariates for 5000 individuals:

\[
z_{1} \sim \textrm{Bernoulli}(0.5) \qquad \qquad %technically a control and treatment group
z_{2} \sim \begin{cases}
 N(0,0.5) & \textrm{if } z_{1}=0\\ 
 N(1,0.5) & \textrm{if } z_{1}=1
\end{cases} \qquad \qquad
z_{3} \sim N(1,0.5).
\]
In addition to the methods described above, we include the exact functional form of the covariates in a CBLR model (referred to as Optimal for simplicity)
in the complex simulation. We obtain confidence intervals by conducting 100 bootstrap re-samples on the training data. We keep 15\% of the data for testing before
hyperparameter selection. We use 15\% of the remaining data for validation and the rest is reserved for training. We predict risk functions for individuals in
the test set, which are used to calculate our IPA and $AUC_{IPCW}$.

@@

Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} shows the performance over time on a test set. The Optimal model acts as a reference for ideal performance on the simulated data. For discrimination, the Optimal model performs best,
followed by CBNN, DeepHit, DeepSurv and the linear models (Figure \ref{fig:megaPlot} E). For both   To obtain a more realistic performance assessment, we compared
models in three case studies with a time-to-event outcome.

@@

\begin{center}\includegraphics[width=1\linewidth]{../figures/Table2.pdf} \end{center}

@@

While we apply a full grid search with three-fold cross-validation for all neural network models, there are an infinite number of untested hyperparameters we did not test. A completely
exhaustive search is computationally infeasible, therefore it is reasonable to expect that there is a set of hyperparameters that is better suited for each model
in each study. However, we test a reasonably large range while accounting for potential over-fitting by including dropout \citep{gulli2017}. We provide the same
options to all models aside from DeepHit, which has a method specific hyperparameter $\alpha$ \citep{lee2018DeepHit}. With these limitations in mind, we
summarize the differences in performance across all tested methods.

@@

Our study was motivated by the lack of easily implemented survival models based on neural networks that can approximate time-varying interactions.
This lead us to repurpose the case-base sampling technique, which has previously been used with logistic regression \citep{hanley2009}, to neural network models for a data driven
approach to time-varying interaction modeling. In our paper, We aim to compare CBNNs with existing methods. We assess the discrimination and calibration
of each method in a simulation with time-varying interactions and a complex baseline hazard, and three case studies. CBNNs outperform all competitors in
the complex simulation and two case studies while maintaining competitive performance in a final case study. Once we perform case-base sampling and
adjust for the sampling bias, we can use a sigmoid activation function to predict our hazard function. Our approach provides an alternative to incorporating
censored individuals, treating survival outcomes as binary ones. Forgoing the requirement of custom loss functions, CBNNs only require the use of standard
components in machine learning libraries (specifically, the add layer to adjust for sampling bias and the sigmoid activation function) after case-base sampling.
Due to the simplicity in its implementation and by extension user experience, CBNNs are both a user-friendly approach to data-driven, single event survival analysis and is
easily extendable to any feed-forward neural network framework.

@@

\appendix
%\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
\bibliography{CBNNManuscript}%

\end{document}

>>>>

%% 
%% Copyright 2007-2020 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version. The latest version of this license is in
%% http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01
%%
%% 
%%
%% $Id: elsarticle-template-num.tex 190 2020-11-23 11:12:32Z rishi $
%%
%%
\documentclass[preprint,12pt]{elsarticle}
\usepackage{hyperref}
 \usepackage{amsmath}
 \usepackage{float}
%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

\journal{Machine learning with applications}

@@

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%% addressline={},
%% city={},
%% postcode={},
%% state={},
%% country={}}
%% \fntext[label3]{}

@@

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \affiliation[label1]{organization={},
%% addressline={},
%% city={},
%% postcode={},
%% state={},
%% country={}}
%%
%% \affiliation[label2]{organization={},
%% addressline={},
%% city={},
%% postcode={},
%% state={},
%% country={}}

\author[inst1]{Jesse Islam}

\affiliation[inst1]{organization={McGill University Department of Quantitative Life Sciences},%Department and Organization
 addressline={805 rue Sherbrooke O}, 
 city={Montréal},
 postcode={H3A 0B9}, 
 state={Quebec},
 country={Canada}}

\author[inst2]{Maxime Turgeon}
\author[inst1,inst3]{Robert Sladek}
\author[inst4]{Sahir Bhatnagar}

\affiliation[inst2]{organization={University of Manitoba Department of Statistics},%Department and Organization
 addressline={50 Sifton Rd}, 
 city={Winnipeg},
 postcode={R3T2N2}, 
 state={Manitoba},
 country={Canada}}

\affiliation[inst3]{organization={McGill University Department of Human Genetics},%Department and Organization
 addressline={805 rue Sherbrooke O}, 
 city={Montréal},
 postcode={H3A 0B9}, 
 state={Quebec},
 country={Canada}}

\affiliation[inst4]{organization={McGill University Department of Biostatistics},%Department and Organization
 addressline={805 rue Sherbrooke O}, 
 city={Montréal},
 postcode={H3A 0B9}, 
 state={Quebec},
 country={Canada}}

\begin{abstract}
%% Text of abstract
Neural network-based survival methods can model data-driven covariate interactions. While these methods can provide better predictive performance than regression-based approaches, not all can model time-varying interactions and complex baseline hazards. To address this, we propose Case-Base Neural Networks (CBNNs) as a new approach that combines the case-base sampling framework with flexible neural network architectures. Using a novel sampling scheme and data augmentation to naturally account for censoring, we construct a feed-forward neural network that may take time as an input. CBNNs predict the probability of an event occurring at a given moment to estimate the hazard function. We compare the performance of CBNNs to regression and neural network-based survival methods in a simulation and three case studies using two time-dependent metrics. First, we examine performance on a simulation involving a complex baseline hazard and time-varying interactions to assess all methods, with CBNN outperforming competitors. Then, we apply all methods to three real data applications, with CBNNs outperforming the competing models in two studies and showing similar performance in the third. Our results highlight the benefit of combining case-base sampling with deep learning to provide a simple and flexible modeling framework for data-driven, time-varying interaction modeling of single event survival outcomes. An R package is available at https://github.com/Jesse-Islam/cbnn.
\end{abstract}

%%Research highlights
\begin{highlights}
\item Case-Base Neural Networks (CBNNs) provide a neural network methodology for survival analysis that leverages a statistical sampling technique, removing the need for complex implementations while providing data-driven flexibility in full hazard function estimation.
\item CBNNs provide an advantage over competing neural network survival models, as it naturally accounts for censoring, uses binary cross-entropy as the loss function and predicts a smooth in-time risk function. 
\item CBNNs outperform the competing models in a simulation, two studies and shows similar performance in the third.
\item Our results highlight the benefit of combining case-base sampling with deep learning to provide a simple and flexible modeling
framework for data-driven, time-varying interaction modeling of single event
survival outcomes.
\end{highlights}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
survival analysis, \sep machine learning \sep case-base \sep neural network
%% PACS codes here, in the form: \PACS code \sep code
%\PACS 0000 \sep 1111
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
%\MSC 0000 \sep 1111
\end{keyword}

\end{frontmatter}

%% \linenumbers

@@

Neural networks provide a data-driven approach to approximating interaction terms that benefit prediction. For example,
DeepSurv is a neural network-based proportional hazards model that implements the Cox partial log-likelihood
as a custom loss function \citep{katzman2018DeepSurv}. This results in a stepwise absolute risk curve that cannot
accommodate time-varying interactions. DeepSurv performs better than Cox proportional hazard models in the Rotterdam and German Breast Cancer Study Group and The Molecular Taxonomy of Breast Cancer International Consortium datasets \citep{katzman2018DeepSurv}. It has also been used to develop personalized prediction models for cancers in a number of studies \citep{ds1} \citep{ds2} \citep{ds3}. Previous work suggests we may change the loss function to address
non-proportional hazards \citep{faraggi1995neural}. In contrast, DeepHit assumes an inverse Gaussian distribution
as the baseline hazard \citep{lee2018DeepHit} and specifies each follow-up time of interest in the model. DeepHit
directly estimates survival curves, rather than deriving a hazard function \citep{lee2018DeepHit}. On real datasets,
DeepHit outperformed DeepSurv \citep{lee2018DeepHit}. We note that DeepSurv and DeepHit both use custom loss
functions \citep{katzman2018DeepSurv} \citep{lee2018DeepHit}. In addition, DeepHit introduces a hyperparameter
weighing its two loss functions (negative log-likelihood and ranking losses) \citep{lee2018DeepHit}. Compared
to these neural network methods, regression-based approaches require prior specification of all interaction terms,
which makes it challenging to model covariate effects that change with time. The current neural network models
provide flexibility at the cost of clarity, while regression models provide clarity at the cost of flexibility.

@@

{\centering \includegraphics[width=1\linewidth]{figures/Figure1.pdf}}

@@

To choose our method-specific hyperparameters, we use the Integrated Brier Score (IBS) \citep{graf1999}, which is based on the
Brier Score (BS) and provides a summarized assessment of performance for each model. We assess the performance of models on
a held out dataset using two metrics: 1) the Index of Prediction Accuracy (IPA) \citep{kattan2018index}; and 2) the Inverse probability
censoring weights-adjusted time-dependent area under the receiver operating characteristic curve ($AUC_{IPCW}$) \citep{auc}. The
IPA score is a metric for both discrimination and calibration, while the $AUC_{IPCW}$ provides a metric for discrimination only.

@@

The IPA is a function of time, based on the BS. The IPA score is given by \begin{align}
\textrm{IPA}(t) &= 1-\frac{BS_{model}(t)}{BS_{null}(t)}, \nonumber
\end{align} where \(BS_{model}(t)\) represents the BS over time \(t\) for the model of interest and \(BS_{null}(t)\) represents
the BS if we use an unadjusted Kaplan-Meier (KM) curve as the prediction for all observations \citep{kattan2018index}. The
IPA score has an upper bound of one, where positive values show an increase in performance over the null model and
negative values show that the null model performs better. As an extension of BS, the IPA score assesses both model calibration
and discrimination and shows how performance changes over follow-up time \citep{graf1999} \citep{kattan2018index}.

\hypertarget{auc}{%
\subsubsection{Inverse probability censoring weights-adjusted time-dependent area under the receiver operating characteristic curve ($AUC_{IPCW}$)}\label{auc}}

The IPCW-adjusted AUC ($AUC_{IPCW}$) is a time-dependent metric that considers censoring \citep{auc}. For a given follow-up time of interest,
\begin{align*}
\resizebox{\textwidth}{!}
{
AUC_{IPCW}(t)=\frac{\sum^{n}_{i=1}\sum^{n}_{j=1} I \left(\widehat{F}(t|X_{i})>\widehat{F}(t|X_{j})\right) \cdot I(T_{i}\leq t,\delta_{i}=1) \cdot \left( 1-I(T_{j}\leq t,\delta_{j}=1) \right) \cdot W_{i}(t) \cdot W_{j}(t)} { \sum^{n}_{i=1}\sum^{n}_{j=1} I(T_{i}\leq t,\delta_{i}=1) \cdot \left( 1-I(T_{j}\leq t,\delta_{j}=1) \right) \cdot W_{i}(t) \cdot W_{j}(t) },
}
\end{align*}
where \begin{align*}
W_{i}(t)=\frac{I(T_{i}\leq t,\delta_{i}=1)}{\widehat{G}(T_{i})} +\frac{I(T_{i}>t)}{\widehat{G}(t)}.
\end{align*} The $AUC_{IPCW}$ measures discrimination \citep{auc}. By examining both IPA and $AUC_{IPCW}$,
we can better understand whether a model performs better in terms of calibration or discrimination.

@@

Neural networks require external parameters that constrain the learning process (hyperparameters). We apply the following hyperparameter
optimization procedure to each method for each study. First, we fix a test set with 15\% of the data which we keep aside during hyperparameter selection.
To determine the best hyperparameters for each neural network method, we use a three-fold cross-validated grid search \citep{gulli2017} on the remaining
data (85\% training, 15\% validation) for the following range of hyperparameters:
\[
Search \ space: \begin{cases}
Learning\ rate \sim \{0.001, 0.01\} \\ 
Dropout \sim \{0.01,0.05, 0.1\} \\
First\ layer\ nodes \sim \{50, 75, 100\} \\
Second\ layer\ nodes \sim \{10,25,50\} \\
Number\ of\ batches \sim \{100, 500\} \\
Activation\ function \sim \{ReLU, Linear\}\\
\alpha \sim \{0, 0.5, 1\}. \ (DeepHit \ only)
\end{cases}
\]

@@

\begin{table}[H]
\caption{Hyperparameters selected after three-fold cross-validated grid search along with the average IBS for each neural network model in the
complex simulation (A); multiple myeloma (MM) case study (B); free light chain (FLC) case study (C); and prostate cancer (Prostate) case study (D).}
\label{tab:wins}
\begin{center}\includegraphics[width=1\linewidth]{figures/Table1.pdf} \end{center}
\end{table}

@@

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{figures/Figure2.pdf}

@@

We simulate data to evaluate the performance of CBNN in comparison with existing regression (Cox, CBLR) and neural network (DeepHit, DeepSurv)
methods. We specify a linear combination of each covariate as the linear predictor in the regression-based methods (Cox, CBLR), which contrasts
with neural network approaches that allow for approximations of non-linear interactions. We simulate data based on a complex baseline hazard with time-varying
interactions and 10\% random censoring. We simulate three covariates for 5000 individuals:

\[
\resizebox{\textwidth}{!}
{
z_{1} \sim \textrm{Bernoulli}(0.5) \qquad \qquad %technically a control and treatment group
z_{2} \sim \begin{cases}
 N(0,0.5) & \textrm{if } z_{1}=0\\ 
 N(1,0.5) & \textrm{if } z_{1}=1
\end{cases} \qquad \qquad
z_{3} \sim N(1,0.5).
}
\]

In addition to the methods described above, we include the exact functional form of the covariates in a CBLR model (referred to as Optimal for simplicity)
in the complex simulation. We obtain confidence intervals by conducting 100 bootstrap re-samples on the training data. We keep 15\% of the data for testing before
hyperparameter selection. We use 15\% of the remaining data for validation and the rest is reserved for training. We predict risk functions for individuals in
the test set, which are used to calculate our IPA and $AUC_{IPCW}$.

@@

Figure \ref{fig:megaPlot} A, E and Table \ref{tab:megaTable} shows the performance over time on a test set. The Optimal model acts as a reference for ideal performance on the simulated data. For discrimination, the Optimal model performs best,
followed by CBNN, DeepHit, DeepSurv and the linear models (Figure \ref{fig:megaPlot} E). To obtain a more realistic performance assessment, we compared
models in three case studies with a time-to-event outcome.

@@

\begin{center}\includegraphics[width=1\linewidth]{figures/Table2.pdf} \end{center}

@@

While we apply a full grid search with three-fold cross-validation for all neural network models, there are an infinite number of untested hyperparameters we did not test. A completely
exhaustive search is computationally infeasible, therefore it is reasonable to expect that there exists a set of hyperparameters that is better suited for each model
in each study. However, we test a reasonably large range while accounting for potential over-fitting by including dropout \citep{gulli2017}. We provide the same
options to all models aside from DeepHit, which has a method specific hyperparameter $\alpha$ \citep{lee2018DeepHit}. With these limitations in mind, we
summarize the differences in performance across all tested methods.

@@

Our study was motivated by the lack of easily implemented survival models based on neural networks that can approximate time-varying interactions.
This lead us to apply the case-base sampling technique, which has previously been used with logistic regression \citep{hanley2009}, to neural network models for a data-driven
approach to time-varying interaction modeling. In our paper, We aim to compare CBNNs with existing methods. We assess the discrimination and calibration
of each method in a simulation with time-varying interactions and a complex baseline hazard, and three case studies. CBNNs outperform all competitors in
the complex simulation and two case studies while maintaining competitive performance in a final case study. Once we perform case-base sampling and
adjust for the sampling bias, we can use a sigmoid activation function to predict our hazard function. Our approach provides an alternative to incorporating
censored individuals, treating survival outcomes as binary ones. Forgoing the requirement of custom loss functions, CBNNs only require the use of standard
components in machine learning libraries (specifically, the add layer to adjust for sampling bias and the sigmoid activation function) after case-base sampling.
Due to the simplicity in its implementation and by extension user experience, CBNNs are both a user-friendly approach to data-driven, single event survival analysis and is
easily extendable to any feed-forward neural network framework.

@@

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
 \bibliographystyle{elsarticle-num} 
 \bibliography{cas-refs}

%% else use the following coding to input the bibitems directly in the
%% TeX file.

% \begin{thebibliography}{00}

% %% \bibitem{label}
% %% Text of bibliographic item

% \bibitem{}

% \end{thebibliography}
\end{document}
\endinput
%%
%% End of file `elsarticle-template-num.tex'.
